---
title: "Instruction Manual"
author: "Tobias Konieczka"
date: "`r Sys.Date()`"
output: pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The covid19_vaccine data set obtained from RamiKrispin on Github as part of their "coronavirus" package was selected for analysis. According to the description, the data set comes from Johns Hopkins Centers for Civic Impact global vaccination data, and is presented in long format by defualt. It can be found [here](https://github.com/RamiKrispin/coronavirus). Due to the sheer breadth of data included in this dataset, functions designed to organize and filter the data needed to be constructed. Using these functions, new questions can be asked:
|   What is the vaccination information in all countries in Africa? 
|   How did the percentage of vaccinated individuals in a country change over a five day period?
|   Which country had the highest number of total vaccinations on a certain day?
|   Is there an association between the rate of vaccination and its position on the globe?
These are just a few of the possible questions that can be raised and answered using 

# KonieTobFin Package

The analysis of the data was made possible and streamlined through the 
development of the **KonieTobFin** package. The functions in the package were 
designed to accomplish a number of data management and analytical tasks. The package
can be installed from my Github and loaded in R using the following code snippet:
```{r, warning=FALSE, results='hide', message=FALSE}
devtools::install_github("konieczkat/KonieTobFin")
library(tidyverse)
library("KonieTobFin")
```

## get_data()

The get_data() Function is used to gather the vaccine dataset from github. It will fetch the coronavirus package
from Github, install it onto the user's system, and attaches them to the R session. Use of this function is required for 
use of the other methods in the package.

```{r, message=FALSE, warning=FALSE}
full_vaccine_data <- get_data()
head(full_vaccine_data)
```
The raw vaccination data is presented in tibble form and contains `r length(full_vaccine_data)` columns, which include date, integer, double, and character types. There are also `r nrow(full_vaccine_data)` rows, as the data contains information from 195 countries between `r min(full_vaccine_data$date)`
and `r max(full_vaccine_data$date)`. There are `r nrow(full_vaccine_data)` observations in the dataset.

## relevise() 
A number of the columns in the dataset represent identifiers that will not be used during analysis. The relevise() function was
designed to remove these unnecessary columns for data management purposes. 
```{r, warning=FALSE, message=FALSE}
data <- relevise(full_vaccine_data)
head(data)
```
Seven of the original 15 rows were removed to streamline the analyses. The columns that remain include date, continent, country, total doses administered, the number of people with at least one dose, the population of the locality, as well as the latitude and longitude for each locality.

## Percentage Calculation

A new column representing the percentage of the population that has received at least one dose can be appended to the modified dataset using the
percent_vaccinated() function. 

```{r, warning=FALSE, message=FALSE}
data <- percent_vaccinated(data)
head(data)
```
Dosing information could be present (as in row 1, 2, 4, and 5 of the above table), have a value of 0 (as in the third row), or missing (as given by NA in row 6). Missing data takes the form on NA throughout the dataset, but the functions have been designed to handle them accordingly.


## Filter Methods

Three tibble filtration methods were developed to subset the data for three different purposes. They relate to the data's presence in spacetime.

### filter_by_continent()

The filter_by_continent() method is used to filter the vaccination data by a specified country of interest. For example,
the function can be used to isolate data from all countries in Africa.

```{r, message=FALSE, warning=FALSE}
Africa <- data %>% filter_by_continent(., "Africa")
head(Africa)
```

### filter_by_continent()

The filter_by_country() method can be used to filter the vaccination data by a specified country of interest. For example,
the function can be used to isolate all vaccination data from Mexico.

```{r, warning=FALSE, message=FALSE}
Mexico <- data %>% filter_by_country(., "Mexico")
head(Mexico)
```

\newpage

### filter_by_date()

The filter_by_date() function can be used to filter the vaccination data through a specified period of time. For example,
the function can be used to isolate vaccination data in Mexico from January 13th, 2021, to January 17th, 2021. 
```{r, warning=FALSE, message=FALSE}
mexicoJan <- Mexico %>% filter_by_date(., "2021-01-13", "2021-01-17")
head(mexicoJan)
```
Over a period of five days, `r max(mexicoJan$people_at_least_one_dose) - min(mexicoJan$people_at_least_one_dose)` Mexicans received at least one dose of the covid vaccine. Using this method in conjunction with the other methods allows the user to identify the global vaccination data on a given day.

```{r, warning=FALSE, message=FALSE}
global_Jan_13_2021 <- data %>% filter_by_date(., "2021-01-13", "2021-01-13")
head(global_Jan_13_2021)
```

\newpage
## visualyze_line()
The visualyze_line() function can be used to plot a line graph representing the change in a variable over time. For example,
using the vaccination info for Mexico from January 13th, 2021, to January 17th, 2021, a graph of the percentage of vaccinated people can be produced.

```{r}
mexicoJan %>% visualyze_line(., .$Percent_Vaccinated) 
```

\newpage
Larger ranges can also be used, and the breaks on the x-axis can be adjusted. Let's take a look at the vaccination percentage in Albania from January 1st, 2022 to December 31st, 2022.

```{r, message=FALSE, warning=FALSE}
data %>% filter_by_country(.,"Albania") %>% filter_by_date(., "2022-01-01", "2022-12-31") %>% visualyze_line(., .$Percent_Vaccinated, dateBreaks = "1 month", dateLabels = "%b")
```
We can see that there was a large spike of vaccinations in Albania in early Febraury 2022 following a plateua of vaccinations in mid to late January.
\newpage

## relation_to_location()
The relation_to_location() function can be used to determine the relationship between vaccination information and position in space. For example,
the relationship between total doses and latitude, total doses and longitude, and total doses and longitude and latitude in every country over the course of the study.

```{r, message=FALSE, warning=FALSE}
data %>% relation_to_location(., .$doses_admin, .$lat, .$long)
```
The output of this call demonstrates that there is a clear association between the number of doses administered in a country and that country's location on Earth.
